---
title: "Building global hexagonal grids"
author: "Petr Keil"
date: "October 26, 2016"
output: 
  html_document: 
    highlight: pygments
    number_sections: yes
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The necessary libraries and stuff

```{r, message=FALSE, warning=FALSE}
library(dggridR)
library(rgdal)
library(ggplot2)

# Shapefile of global administrative areas:
admin <- readOGR(dsn="ADMIN", layer="TM_WORLD_BORDERS")
```

# Create the hexagonal grids 

Construction the dggs objects

```{r}
# for cell area of 629,710.6441 km^2
  dggs4 <- dgconstruct(type = "ISEA3H", res = 4)
# for cell area of 209,903.54803 km^2
  dggs5 <- dgconstruct(type = "ISEA3H", res = 5) 
# for cell area of 69,967.84934
  dggs6 <- dgconstruct(type = "ISEA3H", res = 6) 
```

Create the global grids, unfortunately in the `.klm` format. The resulting
objects are just addresess of where the `.klm` files area saved.

```{r}
  G4 <- dgearthgrid(dggs4, frame = TRUE, wrapcells = TRUE, savegrid = FALSE)
  G5 <- dgearthgrid(dggs5, frame = TRUE, wrapcells = TRUE, savegrid = TRUE)
  G6 <- dgearthgrid(dggs6, frame = TRUE, wrapcells = TRUE, savegrid = TRUE)
```

# Switching to QGIS

Here is a critical step: I was unable to read the `.klm` files neither with `rgdal`,
neither with `dggridR`'s `dg_process_kml`. Maybe it will be possible on Windows,

This is what happens:

```{r, error=TRUE}
  dg_process_kml(G4, frame=TRUE, wrapcells=TRUE)
```

What I did is: **I loaded the `.klm` files to QGIS, and saved them as esri shapefiles**.

# Reading the ESRI shapefiles to R

```{r}
  G4.poly <- readOGR(dsn="GRID", layer="GRIDres4")
  G5.poly <- readOGR(dsn="GRID", layer="GRIDres5")
  G6.poly <- readOGR(dsn="GRID", layer="GRIDres6")
```

# Simple plots

```{r, fig.width=14, fig.height=8}
par(mai=c(0.1, 0.1, 0.1, 0.1))

plot(admin, col="grey", border=NA)
plot(G4.poly, add=TRUE)

plot(admin, col="grey", border=NA)
plot(G5.poly, add=TRUE)

plot(admin, col="grey", border=NA)
plot(G6.poly, add=TRUE)

```

# Fancy ggplots

## Function preparing the shapefile for ggplot2

This funciton thakes the hexagonal grid, and converts it into a data frame
that is usable by `ggplot2`. It also sorts out the ugly wrapping of the hexagons
around the edges.

The function is adopted from the original `dg_process_kml` in package `dggridR`. 
It takes shapefiles instead of the `.klm` files

**Arguments:** 

**map** - `SpatialPolygonsDataFrame` or `SpatialPolygons` class representing the grid. 

```{r}
clean.hex.map <- function(map)
{
  long  <- NULL
  group <-NULL
  
  map@data$timestamp    <- NULL
  map@data$begin        <- NULL
  map@data$end          <- NULL
  map@data$altitudeMode <- NULL
  map@data$extrude      <- NULL
  map@data$visibility   <- NULL
  map@data$drawOrder    <- NULL
  map@data$icon         <- NULL
  map@data$description  <- NULL
  map@data$tessellate   <- NULL
  map@data$id           <- rownames(map@data)
  
  map.points <- fortify(map, region="id")
  map.df     <- merge(map.points, map@data, by="id")
    
  # Find dangerous polygons based on how many degrees of longitude they span
  groups_to_wrap <- map.df %>% group_by(group) %>% 
                    summarise(diff=max(long)-min(long)) %>% 
                    filter(diff>180) %>% select(group)
      
  # Adjust them so they appear on the eastern side of the map
  map.df <- map.df %>% 
            mutate(long=ifelse(group %in% groups_to_wrap$group,
                               ifelse(long<0,long+360,long), long))
  
  # Arrange polygon points so they are ordered appropriately, otherwise the results
  # will not be nice, closed cells, but weird triangular thingies
  clean.map <- map.df %>% arrange(group,order)
  return(clean.map)
}
```

## Processing the shapefiles to data.frames for ggplot2

```{r}
  gg.map4 <- clean.hex.map(G4.poly)
  gg.map5 <- clean.hex.map(G5.poly)
  gg.map6 <- clean.hex.map(G6.poly)
  gg.admin <- fortify(admin)
```

## The plots

```{r, fig.width=14, fig.height=8}
  ggplot(gg.map4, aes(y=lat, x=long, group=group)) +
    theme_bw() +
    geom_polygon(data=gg.admin, aes(y=lat, x=long, group=group), fill="grey") +
    geom_polygon( colour="black", fill=NA)
```

```{r, fig.width=14, fig.height=8}
  ggplot(gg.map5, aes(y=lat, x=long, group=group)) +
    theme_bw() +
    geom_polygon(data=gg.admin, aes(y=lat, x=long, group=group), fill="grey") +
    geom_polygon( colour="black", fill=NA)
```

```{r, fig.width=14, fig.height=8}
  ggplot(gg.map6, aes(y=lat, x=long, group=group)) +
    theme_bw() +
    geom_polygon(data=gg.admin, aes(y=lat, x=long, group=group), fill="grey") +
    geom_polygon( colour="black", fill=NA)
```

# The end

```{r}
  sessionInfo()
```
